### 2020-01-07 qualtrics_forms/instructions.Rmd
minor update

---
title: "Preregistration: data extraction from epidemiology articles"
author: "Katie Drax"
date: "4 December 2019"
output: word_document
---

# Time of writing

- KD and MG independently identifed the study designs (cohort, case-control or cross-sectional) used by each of the 178 "observational epidemiology" articles by completing the Study Design Form for each article. 
In the form, extractors also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
Any conflicts were discussed by the two extractors and resolved by mutual consent if possible. 
If not resolved after discussion, RR determined the study designs. 
The Study Design Form was hosted on Qualtrics but a PDF version can be found here **[]**.
- Data extraction has not yet begun for any eligible articles
- The Data Extraction form has been piloted on 5 observational epidemiology articles by MG, KD and RR. Major changes were made so the Data Extraction Form will be completed again for the piloted articles.
- KD assessed examined the notices attached to the online version of the observational epidemiology articles to identify any retracted articles. None were identified

# Data extraction procedure

The following procedure will be used for all articles classified as 'observational epidemiology'

1. We will attempt to retrieve any supplementary material we could not access (e.g., by contacting the authors). 
1. Articles containing multiple different study designs will be discussed on a case-by-case basis to confirm whether a single reporting checklist is adequate, or if multiple reporting checklists apply to a single article.
1. The information in Table 1 will be extracted manually by KD or automatically by Endnote.
1. The articles will be randomly selected for data extraction and two people (from MG, KD, BW and RR) will independently complete the Data Extraction Form for each article. 
Any conflicts will be discussed by the two extractors and resolved by mutual consent if possible. 
If not resolved after discussion, RR or MM will assign the category. 
Articles will continue to be selected at random until none remain, or until the authors resources are exhausted.

*Table 1. Methods for extracting article meta-data*

| Item | Definition | Method of extraction |
| --- | --- | --- |
| Article ID(s) of related articles | Article IDs of articles that contain similar analyses. | Examine articles with same first author, last author and UK Biobank application number. |
| Citation |   | Endnote |
| Publication date |   | Endnote |
| Authors |   |  Endnote |
| Journal |   | Endnote  |
| Volume |   | Endnote  |
| Issue |   | Endnote  |
| Open Access |   | See if can access article on publishing journal's site on a non-university network. If not accessible search for self-archived version on Google Scholar |
| Countries of corresponding author |   | Extract country from author's affiliations |
| UK Biobank Application number | Study's approved UK Biobank data application number (if present) | Search article for all these terms:  'application' , 'approv' , 'number' , 'Resource'. If not found examine any strings of numbers in article |
| Author keywords |   | Search page 1. If not on page 1 search article for 'keywords' |
| Journal endorsement | Journal's endorsement of STROBE before publication | Extract references to STROBE from the journal's most recent 'instructions/guide for authors' before article's publication date |
| Journal impact factor |   | Retrieve the impact factor of the journal for year of publication from Web of Science's Journal Citation Reports (JCR) |
| Corresponding author email address |   | Endnote; if missing: search article for 'corresponding', '@') |
| Conflicts of interest statement |   | Search article material for all these terms: 'competing', 'interest' ,'conflict' , 'disclosure'grcy |
| Classification | Classification ('mendelian randomisation', 'observational epidemiology' or 'other') as assigned by KD, BW and RR in Rayyan | Rayyan |

# Missing items

We will not attempt to retrieve any missing items in the Data Extraction Form, in order to give an accurate assessment of the version of record and to avoid exhausting the extractors resources.